Theories of deliberated choice

### 2024-04-30 simple.tex
Reordered

  Given a set of attitudes $S$ and $\hist \in \allhists$, define the equivalence relation $=_\hist$ on $S$ as $f =_\hist g$ iff $f(\hist) = g(\hist)$. This also defines the quotient set $S / =_\hist$. (For the readers used to that terminology: if defining $\beta_\hist \in P^S$ as $\beta_\hist(f) = f(\hist)$, then $=_\hist$ is the equivalence kernel of $\beta_\hist$ and $S / =_\hist$ is the coimage of $\beta_\hist$.)
  Similarly, given $S$ and $H \in \finite(\incr(\allhists))$, define the equivalence relation $=_H$ on $S$ as $f =_H g$ iff $\restr{f}{H} = \restr{g}{H}$. This also defines the quotient set $S / =_H$.
\end{definition}
\begin{remark}
  A few more, maybe not useful, definitions.
  A function $l: \allhists \to \powerset{P^{\allhists}}$ is decreasing iff $\forall \hist \wsub \histp: l(\histp) \wsub l(\hist)$.
  Let $S \wsub P^{\allhists}$ be a set of attitudes.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-point-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_\hist$.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-sequence-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_{\intvl{\hist}}$.
\end{remark}
Let $\intvl{\hist} = (\hist_i)_{i \in \intvl{0, \card{\hist}}} \in \finite(\incr(\allhists))$ denote the increasing sequence of paths from $\emptyset$ to $\hist$.
\begin{conjecture}
  The $S$-labeling function associated to an attitude $f \in P^{\allhists}$ is the function $l^S_f: \allhists \to \powerset{P^{\allhists}}$ defined as $l^S_f(\hist) = \set{g \in S \suchthat g(\hist) = f(\hist)}$.

  An attitude is observably included in a set of attitudes $S$ iff $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
\end{conjecture}
\begin{proof}
  Consider $f$ with $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
  Consider any $H \in \finite(\incr(\allhists))$.
  We know that $\cap_{\histp \wsub \hist}{l^S_f(\histp)} \neq \emptyset$, thus in particular, defining $\hist = \max H$ and observing that $H \wsub \set{\histp \suchthat \histp \wsub \hist}$, we have that $\cap_{\histp \in H} l^S_f(\histp) \neq \emptyset$.
  Pick any $g \in \cap_{\histp \in H} l^S_f(\histp)$. Thus, $g \in S$.
  Also, $\forall \hist \in H: g \in l^S_f(\hist)$, thus,
  $\forall \hist \in H: g(\hist) = f(\hist)$.

@@

  Conversely, if $\forall H \in \finite(\incr(\allhists)): \set{g \in S \suchthat \forall \histp \in H: g(\histp) = f(\histp)} \neq \emptyset$, then $\forall \hist \in \allhists$, defining $H = \intvl{\hist}$, we have that $\cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
\end{proof}
\begin{remark}
  The opposite of \cref{th:suffFals} is the negation of $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Thus, for some ${\allleadstog} \in \allleadstos$ and some $i \in I$ with ${\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, we have that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.

  For example, $\allargs = \set{\ar_1, \ar_2}$ and ${\ileadstog}$ is $\phi$ when the path starts with $\ar_1$ (or is empty) and is $¬\phi$ when the path starts with $\ar_2$ and the attitude contains only people constant to $\phi$ (except that they map $\emptyset$ to anything) and people constant to $¬\phi$ (except that they map $\emptyset$ to anything).

  TODO
  In that case, if restricting our focus to claims about types of individuals, we need to include the real types, then close under suitable indistinguishability relation (observable inclusion), so as to include for example $i$.

  Define $S$ as the relations in the claim. Define $R(S)$ as the adequate closure: those that we can’t reject.
  Does $S$ include $R(S)$? If so, it is falsifiable. Otherwise, there is in $R(S)$ a relation that we can’t reject.
  Let $Impl$ be all impliciations in $S$, of the form: if $\hist$ goes to $p$, then $\hist'$ goes to $p'$ (in the above example: if $a$ goes to $\phi$ then $aa$ goes to $\phi$). Any relation that breaks one of these implications is rejectable. For example, sending $a$ to $\phi$ and $ab$ to $¬\phi$ is rejectable.We need to include all relations that break no implication.
  TODO check all this. 
\end{remark}
\begin{lemma}
  \label{th:typesfalsifiable}
  A claim about types of individuals $C^\text{types} \wsub P^{\allhists}$ 
  satisfies $\forall g \in P^{\allhists} \setminus C^\text{types}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in C^\text{types}: \restr{g}{H} \neq \restr{f}{H}$,
  iff its associated claim is falsifiable or trivial.
\end{lemma}
\begin{proof}
  Pick any ${\allleadstog} \notin C$.
  By definition of $C$,
  $\exists i \in I \suchthat {{\ileadstog}} \notin C^\text{types}$.
  As $C$ holds, ${\ileadstog} \notin \set{{\ileadsto[j]} \suchthat j \in I}$ (\cref{th:typesholds}).
  Pick any $H$ such that $\forall {\ileadstog[j]} \in C^\text{types}: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadstog[j]}(\hist)$.
  Define $\theta = \set{(\hist, i, {\ileadstog}(\hist)) \suchthat \hist \in H} \in \Theta \cap \powerset{\allleadstog}$.
  Then, $\theta \notin C^\wsub$, because $\forall {\ileadstog[j]} \in C^\text{types}: \theta \nsubseteq {\ileadstog[j]}$.

@@

\begin{theorem}[To be deleted, probably]
  Consider the attitude where all individuals are either constant to $\phi$ except that they map $\emptyset$ to anything, or constant to $¬\phi$ except that they map $\emptyset$ to anything, say, $(\phi, \phi)$, $(¬\phi, \phi)$, $(\phi, ¬\phi)$, $(¬\phi, ¬\phi)$.
  Consider a claim about types of individuals containing all those ones. Assume that for some attitude $i$ not in the claim, no $H$ permit to observably exclude $i$ from the claim. Then from the lemma above, the claim is not falsifiable. So, we need to include all attitudes that no $H$ permit to observably exclude from the claim. Iterating, this is the closure under the observable inclusion relation.
  So, the claim about types must include the real ones and be closed under observable inclusion.
\end{theorem}
\begin{theorem}
  Consider the attitude above. Is there a claim that holds and is anonymous and falsifiable?
  Check \cref{ex:falsAnonIfBig}.
\end{theorem}

>>>>

  Given a set of attitudes $S$ and $\hist \in \allhists$, define the equivalence relation $=_\hist$ on $S$ as $f =_\hist g$ iff $f(\hist) = g(\hist)$. This also defines the quotient set $S / =_\hist$. (For the readers used to that terminology: if defining $\beta_\hist \in P^S$ as $\beta_\hist(f) = f(\hist)$, then $=_\hist$ is the equivalence kernel of $\beta_\hist$ and $S / =_\hist$ is the coimage of $\beta_\hist$.)
  Similarly, given $S$ and $H \in \finite(\incr(\allhists))$, define the equivalence relation $=_H$ on $S$ as $f =_H g$ iff $\restr{f}{H} = \restr{g}{H}$. This also defines the quotient set $S / =_H$.
\end{definition}
\begin{remark}
  A few more, maybe not useful, definitions.
  A function $l: \allhists \to \powerset{P^{\allhists}}$ is decreasing iff $\forall \hist \wsub \histp: l(\histp) \wsub l(\hist)$.
  Let $S \wsub P^{\allhists}$ be a set of attitudes.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-point-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_\hist$.
  Say that a function $l: \allhists \to \powerset{P^{\allhists}}$ is an $S$-sequence-labeling function iff $\forall \hist \in \allhists: l(\hist) \in S / =_{\intvl{\hist}}$.
\end{remark}
Let $\intvl{\hist} = (\hist_i)_{i \in \intvl{0, \card{\hist}}} \in \finite(\incr(\allhists))$ denote the increasing sequence of paths from $\emptyset$ to $\hist$.
\begin{theorem}
  The $S$-labeling function associated to an attitude $f \in P^{\allhists}$ is the function $l^S_f: \allhists \to \powerset{P^{\allhists}}$ defined as $l^S_f(\hist) = \set{g \in S \suchthat g(\hist) = f(\hist)}$.

  An attitude $f$ is observably included in a set of attitudes $S$ iff $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
\end{theorem}
\begin{proof}
  Consider $f$ with $\forall \hist \in \allhists: \cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
  Consider any $H \in \finite(\incr(\allhists))$.
  We know that $\cap_{\histp \wsub \hist}{l^S_f(\histp)} \neq \emptyset$, thus in particular, defining $\hist = \max H$ and observing that $H \wsub \set{\histp \suchthat \histp \wsub \hist}$, we have that $\cap_{\histp \in H} l^S_f(\histp) \neq \emptyset$.
  Pick any $g \in \cap_{\histp \in H} l^S_f(\histp)$. Thus, $g \in S$.
  Also, $\forall \hist \in H: g \in l^S_f(\hist)$, thus,
  $\forall \hist \in H: g(\hist) = f(\hist)$.

@@

  Conversely, if $\forall H \in \finite(\incr(\allhists)): \set{g \in S \suchthat \forall \histp \in H: g(\histp) = f(\histp)} \neq \emptyset$, then $\forall \hist \in \allhists$, defining $H = \intvl{\hist}$, we have that $\cap_{\histp \wsub \hist} l^S_f(\histp) \neq \emptyset$.
\end{proof}

\begin{lemma}
  \label{th:typesfalsifiable}
  A claim about types of individuals $C^\text{types} \wsub P^{\allhists}$ 
  satisfies $\forall g \in P^{\allhists} \setminus C^\text{types}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall f \in C^\text{types}: \restr{g}{H} \neq \restr{f}{H}$,
  iff its associated claim is falsifiable or trivial.
\end{lemma}
\begin{proof}
  Pick any ${\allleadstog} \notin C$.
  By definition of $C$,
  $\exists i \in I \suchthat {{\ileadstog}} \notin C^\text{types}$.
  As $C$ holds, ${\ileadstog} \notin \set{{\ileadsto[j]} \suchthat j \in I}$ (\cref{th:typesholds}).
  Pick any $H$ such that $\forall {\ileadstog[j]} \in C^\text{types}: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadstog[j]}(\hist)$.
  Define $\theta = \set{(\hist, i, {\ileadstog}(\hist)) \suchthat \hist \in H} \in \Theta \cap \powerset{\allleadstog}$.
  Then, $\theta \notin C^\wsub$, because $\forall {\ileadstog[j]} \in C^\text{types}: \theta \nsubseteq {\ileadstog[j]}$.

@@

\begin{remark}
  The opposite of \cref{th:suffFals} is the negation of $\forall {\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}: \exists H \in \finite(\incr(\allhists)) \suchthat \forall j \in I: \exists \hist \in H \suchthat {\ileadstog}(\hist) \neq {\ileadsto[j]}(\hist)$.
  Thus, for some ${\allleadstog} \in \allleadstos$ and some $i \in I$ with ${\ileadstog} \in P^{\allhists} \setminus \set{{\ileadsto[k]} \suchthat k \in I}$, we have that $\forall H \in \finite(\incr(\allhists)): \exists j \in I \suchthat \forall \hist \in H: {\ileadstog}(\hist) = {\ileadsto[j]}(\hist)$.

  For example, $\allargs = \set{\ar_1, \ar_2}$ and ${\ileadstog}$ is $\phi$ when the path starts with $\ar_1$ (or is empty) and is $¬\phi$ when the path starts with $\ar_2$ and the attitude contains only people constant to $\phi$ (except that they map $\emptyset$ to anything) and people constant to $¬\phi$ (except that they map $\emptyset$ to anything), say, $(\phi, \phi)$, $(¬\phi, \phi)$, $(\phi, ¬\phi)$, $(¬\phi, ¬\phi)$.

  In that case, if restricting our focus to claims about types of individuals, we need to include the real types, then close under suitable indistinguishability relation (observable inclusion), so as to include for example $i$.

  Define $S$ as the relations in the claim. Define $R(S)$ as the adequate closure: those that we can’t reject.
  Does $S$ include $R(S)$? If so, it is falsifiable. Otherwise, there is in $R(S)$ a relation that we can’t reject.
  Let $Impl$ be all impliciations in $S$, of the form: if $\hist$ goes to $p$, then $\hist'$ goes to $p'$ (in the above example: if $a$ goes to $\phi$ then $aa$ goes to $\phi$). Any relation that breaks one of these implications is rejectable. For example, sending $a$ to $\phi$ and $ab$ to $¬\phi$ is rejectable.We need to include all relations that break no implication.
  TODO check all this. 
\end{remark}

\begin{theorem}
  Consider the attitude above. Is there a claim that holds and is anonymous and falsifiable?
  Check \cref{ex:falsAnonIfBig}.
\end{theorem}
